---
title: "Exam 6"
author: "Alice, Kseniia, ShanShan"
date: "02 06 2026"
output:
  html_document:
    toc: true
    toc_depth: 1
    number_sections: true
  pdf_document:
    toc: true
    toc_depth: '1'
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Environment setting
```{r}
#plotting for all task
if (!require("ggplot2")) { install.packages("ggplot2"); library("ggplot2") }

#imputation data for task5 
if (!require("mice")) { install.packages("mice"); library("mice") } 

#different logistic models comparison for task5
if (!require("sjPlot")) { install.packages("sjPlot"); library("sjPlot") } 

#Variance Inflation Factor (VIF) diagnostics for task5
if (!require("car")) { install.packages("car"); library("car") } 

#regression summaries for task5
if (!require("gtsummary")) { install.packages("gtsummary"); library("gtsummary") }

#ROC and AUC analysis for task5
if (!require("pROC")) { install.packages("pROC"); library("pROC") }

#machine learning for task5
if (!require("randomForest")) { install.packages("randomForest"); library("randomForest") }
 
```

# Task 1
1. Load and clean the data sets. Give an overview over the different variables analyze them descriptively.
2. Analyze the differences between the genders descriptively.
```{r task1}

# 1.1 load and clean the datasets
dat1 <- read.csv("Exam_6_1.csv")
summary(dat1)

# uppercase all names
dat1$Name <- toupper(dat1$Name)
# fix the comma instead of dot issue
dat1$Billing.Amount <- gsub(",", ".", dat1$Billing.Amount)
# set negative values of billing amount as NA
dat1$Billing.Amount[dat1$Billing.Amount < 0] = NA

# omit first column
dat1 <- dat1[, -1]

# find and remove duplicates
sum(duplicated(dat1))
dat1 <- dat1[!duplicated(dat1), ]

# fix variable type
dat1$Date.of.Admission <- as.Date(dat1$Date.of.Admission, "%Y-%m-%d")
dat1$Discharge.Date <- as.Date(dat1$Discharge.Date, "%Y-%m-%d")
dat1$Admission.Type <- as.factor(dat1$Admission.Type)
dat1$Blood.Type <- as.factor(dat1$Blood.Type)
dat1$Gender <- as.factor(dat1$Gender)
dat1$Doctor <- as.factor(dat1$Doctor)
dat1$Name <- as.factor(dat1$Name)
dat1$Insurance.Provider <- as.factor(dat1$Insurance.Provider)
dat1$Medical.Condition <- as.factor(dat1$Medical.Condition)
dat1$Test.Results <- as.factor(dat1$Test.Results)
dat1$Room.Number <- as.factor(dat1$Room.Number)
dat1$Billing.Amount <- as.numeric(dat1$Billing.Amount)
dat1$Medication <- as.factor(dat1$Medication)

# calculate duration of hospitalization
dat1$Duration.of.Hospitalization <- as.numeric(dat1$Discharge.Date - dat1$Date.of.Admission)

# remove all missing values
clean_dat1 <- na.omit(dat1) 
summary(clean_dat1)

# load and clean data2
dat2 <- read.csv("Exam_6_2.csv")
summary(dat2)

# fix variable types
dat2$patient_id <- as.factor(dat2$patient_id)
dat2$gender <- as.factor(dat2$gender)
dat2$ace_inhibitor <- as.factor(dat2$ace_inhibitor)
dat2$beta_blocker <- as.factor(dat2$beta_blocker)
dat2$diuretic <- as.factor(dat2$diuretic)
dat2$income_level <- as.factor(dat2$income_level)
dat2$readmitted_30d <- as.factor(dat2$readmitted_30d)

# remove values that make no physiological sense
dat2$creatinine[dat2$creatinine < 0] = NA
dat2$bmi[dat2$bmi < 12] = NA

# find and remove duplicates
sum(duplicated(dat2))
dat2 <- dat2[!duplicated(dat2), ]

# we decided to delete missing data 
clean_dat2 <- na.omit(dat2) 
summary(clean_dat2)

# 1.2 Analyze differences between the genders descriptively
by(clean_dat1, clean_dat1$Gender, summary)
by(clean_dat2, clean_dat2$gender, summary)

# make the plots for gender differences
# gender differences in blood type

ggplot(clean_dat1, aes(x = Blood.Type, fill = Gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Blood Type Distribution by Gender",
       subtitle = "Analysis of baseline patient demographics",
       x = "Blood Type", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences in medical condition
ggplot(clean_dat1, aes(x = Medical.Condition, fill = Gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Medical Condition Distribution by Gender",
       x = "Medical Condition", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences in insurance provider
ggplot(clean_dat1, aes(x = Insurance.Provider, fill = Gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Insurance Provider Distribution by Gender",
       x = "Insurance Provider", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences in billing amount
ggplot(clean_dat1, aes(x = Gender, y = Billing.Amount, fill = Gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Billing Amount Distribution by Gender",
       x = "Gender", y = "Billing Amount ($)") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences in admission type
ggplot(clean_dat1, aes(x = Admission.Type, fill = Gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Admission Type Distribution by Gender",
       x = "Admission Type", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences in Medication
ggplot(clean_dat1, aes(x = Medication, fill = Gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Medication Distribution by Gender",
       x = "Medication", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences in Test Results
ggplot(clean_dat1, aes(x = Test.Results, fill = Gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Test Result Distribution by Gender",
       x = "Test Results", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences in duration of Hospitalization
ggplot(clean_dat1, aes(x = Gender, y = Duration.of.Hospitalization, fill = Gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Duration of Hospitalization Distribution by Gender",
       x = "Gender", y = "Duration of Hospitalization (days)") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences dat2
ggplot(clean_dat2, aes(x = gender, y = bmi, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Body Mass Index Distribution by Gender",
       x = "Gender", y = "BMI") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences bnp
ggplot(clean_dat2, aes(x = gender, y = bnp, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "BNP Distribution by Gender", x = "Gender", y = "BNP") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences sodium
ggplot(clean_dat2, aes(x = gender, y = sodium, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Sodium Distribution by Gender", x = "Gender", y = "Sodium") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences creatinine
ggplot(clean_dat2, aes(x = gender, y = creatinine, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Creatinine Distribution by Gender", x = "Gender", y = "Creatinine") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences systolic_bp
ggplot(clean_dat2, aes(x = gender, y = systolic_bp, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Systolic Blood Pressure Distribution by Gender", x = "Gender", y = "Systolic Blood Pressure (mmHg)") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences heart rate
ggplot(clean_dat2, aes(x = gender, y = heart_rate, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Heart Rate Distribution by Gender", x = "Gender", y = "Heart Rate (BPM)") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences ace inhibitor
ggplot(clean_dat2, aes(x = ace_inhibitor, fill = gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "ACE Inhibitor Use Distribution by Gender", x = "ACE Inhibitor Use [1 (YES); 0 (NO)]", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences beta blocker
ggplot(clean_dat2, aes(x = beta_blocker, fill = gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Beta Blocker Use Distribution by Gender", x = "Beta Blocker Use [1 (YES); 0 (NO)]", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences diuretic
ggplot(clean_dat2, aes(x = diuretic, fill = gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Diuretic Use Distribution by Gender", x = "Diuretic Use [1 (YES); 0 (NO)]", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences adherence score
ggplot(clean_dat2, aes(x = gender, y = adherence_score, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Adherence score Distribution by Gender", x = "Gender", y = "Adherence score") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences income level
ggplot(clean_dat2, aes(x = income_level, fill = gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Income Level Distribution by Gender", x = "Income Level", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

# gender differences distance to hospital
ggplot(clean_dat2, aes(x = gender, y = distance_to_hospital_km, fill = gender)) +
  geom_boxplot(width = 0.5, outlier.alpha = 0.4) +
  theme_bw() + 
  labs(title = "Distance to Hospital Distribution by Gender", x = "Distance to Hospital (km)", y = "Distance (km)") +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# gender differences readmitted 30 days
ggplot(clean_dat2, aes(x = readmitted_30d, fill = gender)) +
  geom_bar(position = "dodge") +
  theme_bw() + 
  labs(title = "Patients readmitted after 30 days Distribution by Gender", x = "Readmitted after 30 days [1 (YES); 0 (NO)]", y = "Patient Count") +
  theme(plot.title = element_text(face = "bold"))

```

# Task 2
The following tasks refer to: Exam data 6 1.csv
Investigate the effect of the following variables on the billing amount:
1. Date of admission
2. Duration of stay
3. Medical condition
```{r task2}

clean_dat1$Admission_Year <- as.factor(format(as.Date(clean_dat1$Date.of.Admission), "%Y"))

# Check the skewness of billing
hist(clean_dat1$Billing.Amount, 
     probability = TRUE, 
     breaks = 50, 
     main = "Distribution of Billing Amounts", 
     xlab = "Total Billing Amount", 
     col = "lightgray")
lines(density(clean_dat1$Billing.Amount, na.rm = TRUE), col = "red",  lwd = 2)


# First visualize the data
par(mfrow = c(1, 3))

plot(Billing.Amount ~ Admission_Year, data = clean_dat1, 
     main = "Billing by Year", col = "#E87D72")
plot(Billing.Amount ~ Duration.of.Hospitalization, data = clean_dat1, 
     main = "Billing by Duration", col = "#54B84C")
plot(Billing.Amount ~ Medical.Condition, data = clean_dat1, 
     main = "Billing by Condition", col = "#6F9BF8")

par(mfrow = c(1, 1))

# Second run bivariate models

# 1. Effect of Time alone

# Run the updated bivariate model
model_year <- lm(Billing.Amount ~ Admission_Year, data = clean_dat1)
summary(model_year)

# 2. Effect of Hospitalization Duration alone
model_duration <- lm(Billing.Amount ~ Duration.of.Hospitalization, data = clean_dat1)
summary(model_duration)

# 3. Effect of Medical Condition alone
model_condition <- lm(Billing.Amount ~ Medical.Condition, data = clean_dat1)
summary(model_condition)

# Then combined model
model_combined <- lm(Billing.Amount ~ Admission_Year + Duration.of.Hospitalization + Medical.Condition, data = clean_dat1)
summary(model_combined)

```

# Task 3
The following tasks refer to: Exam data 6 1.csv
1. For this task remove all observations with inconclusive test results.
2. Fit a logistic regression model with test result as the dependent outcome. You are free to choose which
independent variables to include.
```{r task3}

# 3.1 Remove all observations with inconclusive test results
dat1_conclusive <- clean_dat1[clean_dat1$Test.Results != "Inconclusive", ]
dat1_conclusive$Test.Results <- droplevels(dat1_conclusive$Test.Results)

# Check remaining levels
table(dat1_conclusive$Test.Results)

# 3.2  Logistic regression with the test result as outcome
# Create binary outcome: 1 = Abnormal, 0 = Normal
dat1_conclusive$Test_binary <- 0
dat1_conclusive[dat1_conclusive$Test.Results == "Abnormal", "Test_binary"] <- 1

# Fit logistic regression: use age, gender, medical condition, admission type, duration
mod_logit <- glm(Test_binary ~ Age + Gender + Medical.Condition +
                   Admission.Type + Duration.of.Hospitalization,
                 data = dat1_conclusive,
                 family = binomial)
summary(mod_logit)

# Alternative model with fewer predictors
mod_logit2 <- glm(Test_binary ~ Age + Medical.Condition + Duration.of.Hospitalization,
                  data = dat1_conclusive,
                  family = binomial)
summary(mod_logit2)

# Compare AIC of both models
AIC(mod_logit, mod_logit2)
anova(mod_logit2, mod_logit, test = "LRT") 

```

# Task 4
The following tasks refer to: Exam data 6 2.csv
Income analysis:
1. Is there an association between the income level and the distance to the hospital?
2. Are there any clinical variables which differ significantly between income levels?
3. How is age distributed in each income level?
```{r task4}
# 4.1 Association between income level and distance to the hospital?
ggplot(clean_dat2, aes(x = income_level, y = distance_to_hospital_km, fill = income_level)) +
  geom_boxplot() +
  ggtitle("Distance to Hospital by Income Level") +
  theme_bw()

# Mean distance per income level
tapply(clean_dat2$distance_to_hospital_km, clean_dat2$income_level, mean)

# income_level has 3 groups, so we use ANOVA
anova_dist <- aov(distance_to_hospital_km ~ income_level, data = clean_dat2)
summary(anova_dist)

# 4.2  Are there any clinical variables which differ significantly between income levels?
summary(aov(bmi             ~ income_level, data = clean_dat2))
summary(aov(bnp             ~ income_level, data = clean_dat2))
summary(aov(sodium          ~ income_level, data = clean_dat2))
summary(aov(creatinine      ~ income_level, data = clean_dat2))
summary(aov(systolic_bp     ~ income_level, data = clean_dat2))
summary(aov(heart_rate      ~ income_level, data = clean_dat2))
summary(aov(adherence_score ~ income_level, data = clean_dat2))

# Nothing is significant, lets just visualize one with lower p-value
ggplot(clean_dat2, aes(x = income_level, y = systolic_bp, fill = income_level)) +
  geom_boxplot() +
  ggtitle("Systolic BP by Income Level") +
  theme_bw()

# 4.3 How is age distributed in each income level?
ggplot(clean_dat2, aes(x = age, fill = income_level)) +
  geom_histogram(binwidth = 5, color = "white") +
  facet_wrap(~ income_level) +
  ggtitle("Age Distribution by Income Level") +
  theme_bw()

# Summary statistics
tapply(clean_dat2$age, clean_dat2$income_level, summary)

```

# Task 5
The following tasks refer to: Exam data 6 2.csv
Readmission prediction:
1. Try to predict which patients are readmitted to the hospital.
2. Try different model specifications and discuss them.
3. Which variables are most important?
```{r task5}
colSums(is.na(dat2)) # 92 NAs in bmi, 90 NAs in sodium, 108 NAs in creatinine
dat2_lossall <- dat2[is.na(dat2$bmi) & is.na(dat2$sodium) & is.na(dat2$creatinine),] 
nrow(dat2_lossall)
dat2_loss2 <- dat2[is.na(dat2$creatinine) & is.na(dat2$sodium),]
nrow(dat2_loss2)
# bmi&c: 0; sodium&c: 5;  sodium&bmi: 5; bmi&c&sodium: 0; total missing: 280 

# create dataset for machine learning
## drop patient_id
dat2_ml <- dat2[,-1]
# Order the income factor levels properly
dat2_ml$income_level <- factor(dat2_ml$income_level,
                                  levels = c("Low", "Medium", "High"),
                                  ordered = TRUE)

# Train/Test split (80/20)
set.seed(500)
sample_size <- floor(0.8 * nrow(dat2_ml))
train_indices <- sample(seq_len(nrow(dat2_ml)), size = sample_size)

# impute the missing data
ignore <- rep(FALSE, nrow(dat2_ml))
ignore[-train_indices] <- TRUE
set.seed(123)
imputation <- mice(dat2_ml, m = 1, method = 'pmm', 
                   ignore = ignore, printFlag = FALSE)
dat2_ml <- complete(imputation, 1)
train_dat <- dat2_ml[train_indices, ]
test_dat <- dat2_ml[-train_indices, ]

# First Model : Logistic Regression

# Checking multicollinearity
num_vars <- train_dat[, c("age", "distance_to_hospital_km", "bmi", "bnp", 
                          "sodium", "creatinine", "systolic_bp", "heart_rate", 
                          "adherence_score")]
cor_matrix <- cor(num_vars, use = "complete.obs")
round(cor_matrix, 2) # all look safe

# Stepwise Selection
model_all <- glm(readmitted_30d ~ ., data = train_dat, family = binomial)
model_sele <- step(model_all, direction = "both", trace = FALSE)
summary(model_sele)

# Hierarchical Models & ANOVA Block Testing
# Block 1: Baseline Characteristics
model1 <- glm(readmitted_30d ~ age, data = train_dat, family = binomial) #AIC:3227.6 
model2 <- glm(readmitted_30d ~ age + distance_to_hospital_km, data = train_dat, family = binomial) #AIC:3220.2
anova(model1, model2, test = "LRT") 

# Block 2: Add Labs (physiological markers)
model3 <- glm(readmitted_30d ~ age + distance_to_hospital_km + bnp + sodium +
                creatinine, data = train_dat, family = binomial) #AIC:3152.1
anova(model2, model3, test = "LRT") 

model3.1 <- glm(readmitted_30d ~ age + distance_to_hospital_km + bnp + sodium + creatinine +
                systolic_bp + heart_rate, 
              data = train_dat, family = binomial) #AIC: 3148.2
anova(model3, model3.1, test = "LRT") #choose 3.1

# Block 3: Add Medications
model4 <- glm(readmitted_30d ~ age + distance_to_hospital_km + bnp + sodium + creatinine +
                systolic_bp + heart_rate + ace_inhibitor + beta_blocker,
                data = train_dat, family = binomial) #AIC:3113.9
anova(model3.1, model4, test = "LRT") 

# Block 4: Add Behavior
model5 <- glm(readmitted_30d ~ age + distance_to_hospital_km + bnp + sodium + creatinine +
                systolic_bp + heart_rate +
                ace_inhibitor + beta_blocker + adherence_score, 
                data = train_dat, family = binomial) #AIC:3037.1
anova(model4, model5, test = "LRT") 

#drop systolic bp and heart rate
model5.1 <- glm(readmitted_30d ~ age + distance_to_hospital_km + bnp + sodium + creatinine +
                ace_inhibitor + beta_blocker + adherence_score, 
                data = train_dat, family = binomial) #AIC:3038.9

anova(model5.1, model5, test = "LRT") 
#model5 doesn't significant improve the model, choose model 5.1

# Interaction checks (both were non-significant >> choose model5)

# Case A: Does patient age modify the effect of medication adherence?
model_int_age <- update(model5.1, . ~ . + age:adherence_score) #AIC:3040.8
anova(model5.1, model_int_age, test = "LRT") 

# Case B: Does diuretic use alter the risk context of blood sodium levels?
model_int_sod <- update(model5.1, . ~ . + diuretic:sodium) #AIC:3039.3
anova(model5.1, model_int_sod, test = "LRT") 

# Model Outputs & Tables
tab_model(model2, model3.1, model4,  model5, model5.1, model_int_age, model_int_sod,
           show.ci = FALSE, show.aic = TRUE, p.style = "numeric_stars")

vif(model5.1) # check multicollinearity (all variance inflation factor near 1.0)


# the final one is model 5.1
model_lr <- model5.1

# Second Model: Random Forest
set.seed(42)
#default model
model500 <- randomForest(readmitted_30d ~ ., data = train_dat, proximity=FALSE)

oob.error.data500 <- data.frame(
  Trees=rep(1:nrow(model500$err.rate), times=3),
  Type=rep(c("OOB", "notreadmitted", "readmitted"), each=nrow(model500$err.rate)),
  Error=c(model500$err.rate[,"OOB"], 
          model500$err.rate[,"0"], 
          model500$err.rate[,"1"]))

ggplot(data = oob.error.data500, aes(x = Trees, y = Error)) +
  geom_line(aes(color = Type)) +
  theme_bw() +
  labs(title = "OOB Error Convergence Profile (500 Trees)")

#check how many tree is better
model1000 <- randomForest(readmitted_30d ~ ., data = train_dat,
                          ntree = 1000, proximity=FALSE) 
oob.error.data1000 <- data.frame(
  Trees=rep(1:nrow(model1000$err.rate), times=3),
  Type=rep(c("OOB", "notreadmitted", "readmitted"), each=nrow(model1000$err.rate)),
  Error=c(model1000$err.rate[,"OOB"], 
          model1000$err.rate[,"0"], 
          model1000$err.rate[,"1"]))

ggplot(data = oob.error.data1000, aes(x = Trees, y = Error)) +
  geom_line(aes(color = Type)) +
  theme_bw() +
  labs(title = "OOB Error Convergence Profile (1000 Trees)") # no significant leap >> keep default tree

# Tune mtry
max_pred <- ncol(train_dat)-1
oob.values <- vector(length=max_pred) 

set.seed(233)
for(i in 1:max_pred) {
  temp.model <- randomForest(readmitted_30d ~ ., mtry=i,
                             data = train_dat, proximity=FALSE)
  oob.values[i] <- temp.model$err.rate[nrow(temp.model$err.rate), "OOB"]
}
oob.values #again no big difference >> use default setting

# Final Model Definition
set.seed(42)
model_rf <- randomForest(readmitted_30d ~ . ,data = train_dat, proximity=FALSE)

# Model Evaluation on Test Set
log_preds <- predict(model_lr, newdata = test_dat, type = "response")
rf_preds <- predict(model_rf, newdata = test_dat, type = "prob")[, "1"]

# Optimize threshold
log_roc <- roc(test_dat$readmitted_30d, log_preds, quiet = TRUE, percent = TRUE)
rf_roc <- roc(test_dat$readmitted_30d, rf_preds, quiet = TRUE, percent = TRUE)
log_opt <- coords(log_roc, "best", ret = c("threshold", "specificity", "sensitivity"), best.method = "youden")
rf_opt <- coords(rf_roc, "best", ret = c("threshold", "specificity", "sensitivity"), best.method = "youden")

# Using clinical cutoff threshold of 0.40
log_outcome <- ifelse(log_preds > 0.4, 1, 0)
rf_outcome <- ifelse(rf_preds > 0.4, 1, 0)
log_comp_tab <- table(Predicted = log_outcome, Actual = test_dat$readmitted_30d)
rf_comp_tab <- table(Predicted = rf_outcome, Actual = test_dat$readmitted_30d)

# Compute final metrics
log_t_positive <- log_comp_tab["1", "1"]
log_t_negative <- log_comp_tab["0", "0"]
log_f_positive <- log_comp_tab["1", "0"]
log_f_negative <- log_comp_tab["0", "1"]

log_accuracy    <- (log_t_positive + log_t_negative) / sum(log_comp_tab)
log_sensitivity <- log_t_positive / (log_t_positive + log_f_negative)
log_specificity <- log_t_negative / (log_t_negative + log_f_positive)

rf_t_positive <- rf_comp_tab["1", "1"]
rf_t_negative <- rf_comp_tab["0", "0"]
rf_f_positive <- rf_comp_tab["1", "0"]
rf_f_negative <- rf_comp_tab["0", "1"]

rf_accuracy    <- (rf_t_positive + rf_t_negative) / sum(rf_comp_tab)
rf_sensitivity <- rf_t_positive / (rf_t_positive + rf_f_negative)
rf_specificity <- rf_t_negative / (rf_t_negative + rf_f_positive)

# Comparison
cat("\n--- Logistic Regression PERFORMANCE (THRESHOLD = 0.4) ---\n",
    "Accuracy    : ", round(log_accuracy, 4), "\n",
    "Sensitivity : ", round(log_sensitivity, 4), "\n",
    "Specificity : ", round(log_specificity, 4), "\n")
cat("\n--- RANDOM FOREST PERFORMANCE (THRESHOLD = 0.4) ---\n",
    "Accuracy    : ", round(rf_accuracy, 4), "\n",
    "Sensitivity : ", round(rf_sensitivity, 4), "\n",
    "Specificity : ", round(rf_specificity, 4), "\n")

# ROC Curve & AUC
ggroc(list(Logistic = log_roc, `Random Forest` = rf_roc), legacy.axes = TRUE, linewidth = 1.2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  theme_bw() +
  scale_color_manual(values = c("Logistic" = "#377eb8", "Random Forest" = "#e41a1c")) +
  labs(title = "ROC Curve Comparison",
       subtitle = paste0("Logistic AUC = ", round(log_roc$auc, 2), 
                         " | Random Forest AUC = ", round(rf_roc$auc, 2)),
       x = "False Positive Percentage (1 - Specificity)",
       y = "True Positive Percentage (Sensitivity)",
       color = "Model Specification")

# Variable Importance Plot (OR and absolute z-values)
tbl_regression(model_lr, exponentiate = TRUE) #creat summary table

coef_mod <- summary(model_lr)$coefficients[-1, ]
z_df <- data.frame(
  Variable = rownames(coef_mod),
  Importance = abs(coef_mod[, "z value"])
)

ggplot(z_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_col(fill = "#377eb8", width = 0.6) +                  
  geom_text(aes(label = round(Importance, 2)),hjust = -0.2, size = 4) +
  coord_flip() +
  theme_bw() +                               
  labs(title = "Variable Importance (Absolute z-values)", x = NULL, y = "z-value") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))

```




